---
title: "Markdown & Literate Coding"
output: html_notebook
---

## Markdown

A [lightweight markup language](https://en.wikipedia.org/wiki/Lightweight_markup_language) for *creating formatted text* using a **plain-text editor**. John Gruber and Aaron Swartz created Markdown in 2004 as a markup language that is appealing to human readers in its source code form. ^[https://en.wikipedia.org/wiki/Markdown]

## Literate Coding

A technique where coding is interspersed with natural language. Used in support of reproducible research ^[https://en.wikipedia.org/wiki/Literate_programming] 

Examples include

-   R Markdown
-   Jupyter Notebooks

## Code chunks

This is a code chunk

```{r example, eval=FALSE}
# nothing to see here
```

... with code inside it

```{r plot cars}
plot(cars)
```

The code chunks can be hidden or displayed:

-   show output only
-   show code and output
-   show nothing run code
-   show nothing (dont' run code)

### Get Started

> `File > New File > R Notebook`

Every time you open a new R Notebook, the document will be pre-formatted with example text and code that can be used to learn how to use R Markdown and literate codding.

## Visual Editor

Starting with RStudio version 1.4.x you can use a visual (WYSIWYG) editor to edit R markdown documents. But not Xaringan.

## Output formats

There are multiple output formats that can be generated from `rmarkdown::render()`. Those format types are identified in the `output` element of the YAML header -- used for report configuration.

Some format options include

-   html_notebooks (while you develop your script)
-   html_document (for final production)
-   github_document (for github)
-   MS Word file
-   PDF document (requires some LaTEX engine - a mild pain to configure)
-   Dashboards
-   Slides (including Xaringan)
-   ebooks (bookdown)
-   web sites (blogdown, Distill)

## YAML configuration

The YAML header is at the top of the document. When you first opened this document the YAML header started at line 1 (always) and ends at line 4 with three dashes:  `---`. (context specific.)

## Render reports

Most of the time you'll render your report by using he Knit button (or the Preview button) at the top of the script editor. But you can use the `rmarkdown::render()` function from the console.

## Help

There are three help documents for R Markdown avaialble from the Help menu of RStudio, Two are in the *cheatsheets* sub-section, the other is a top level within Help. Also [this printable cheatsheet](https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf)

See Also:

-   R Markdown - <https://rmarkdown.rstudio.com/>
-   *R Markdown: The Definitive Guide* - <https://bookdown.org/yihui/rmarkdown/>
